Illusory Finger Stretching and Somatosensory Responses

### 2024-08-28 manuscript.qmd
Update manuscript.qmd
Script updated to reflect revision changes for stage 2 version 2

---
title: "Somatosensory Response Changes During Illusory Finger Stretching"
author: ""
bibliography: references-rr.bib
#csl: elife.csl
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: true
execute: 
  echo: false
  include: false
  output: false
format: 
  pdf:
    fig-pos: "H"
---

@@

# install R packages
packagelist <- c('tidyverse','ggpubr','ez','ggplot2','jpeg','grid','plotrix','lavaan', 'cowplot', 'ggpubr','rstatix','reticulate','moments','tiff','rmarkdown','tinytex','osfr')  
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

pdf('Figures/eegdata1.pdf', width = 12, height = 11)
#visualise data
  ggplot(eegdata_2, aes(x = Condition, y = Amplitude)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("SSEP Amplitude (µV)")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_y_log10() +
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  #geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               #c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               #c("UV.Illusion.Median","MS.Illusion.Median")),
              #map_signif_level = TRUE,annotations="***",y_position = c(115,135,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

pdf('Figures/eegdata2.pdf', width = 12, height = 11)
#visualise data
  ggplot(eegdata_exp2, aes(x = Condition, y = Amplitude)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("SSEP Amplitude (µV)")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_y_log10() +
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  #geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               #c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               #c("UV.Illusion.Median","MS.Illusion.Median")),
              #map_signif_level = TRUE,annotations="***",y_position = c(115,135,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#get medians and sds for eeg data
eegdatamsds <- group_by(eegdata_2, Condition) %>%
  summarise(
    count = n(),
    median = median(Amplitude, na.rm = TRUE),
    sd = sd(Amplitude, na.rm = TRUE)
  )

@@

#run one way RMANOVA for illusion data
#res.aov4 <- anova_test(data = eegdata_2, dv = Amplitude, wid = PID, within = Condition)
#get_anova_table(res.aov4)

#posthoc tests
# pwc4 <- eegdata_2 %>%
# pairwise_t_test(
# Amplitude ~ Condition, paired = TRUE,
# p.adjust.method = "bonferroni"
# )
# pwc4

@@

#posthoc tests
nppwc_eeg1 <- eegdata_2 %>%
  wilcox_test(Amplitude ~ Condition, paired = TRUE, comparisons = list(c("V1", "V3"), c("V1", "V4"), c("V1", "V2")), p.adjust.method = "holm")

nppwc2_eeg2 <- eegdata_exp2 %>%
  wilcox_test(Amplitude ~ Condition, paired = TRUE, comparisons = list(c("V1", "V3"), c("V1", "V4"), c("V1", "V2")), p.adjust.method = "holm")

@@

#create spectra plots
pdf(paste0(figdir,"/spectrum1.pdf"), width=12, height=12)
par(mfrow=c(2,2))
# plot the Fourier spectra across all electrodes
f <- seq(0,499.5,0.5)

@@

polygon(f[c(3:100,100:3)],c(meanfspec[cond,3:100]+SEfspec[cond,3:100],meanfspec[cond,100:3]-SEfspec[cond,100:3]),col=addalpha('cornflowerblue',0.25),border=NA)
lines(f[3:100],meanfspec[cond,3:100],lwd=3) 
# lines(c(0,50), meanfspec[1,c(53,53)],lty=2)
# lines(c(26,26),c(0,0.5),col='red')
}
dev.off()

#create exploratory spectra plots
pdf(paste0(figdir,"/spectrum2.pdf"), width=12, height=12)
par(mfrow=c(2,2))
# plot the Fourier spectra across all electrodes
f <- seq(0,499.5,0.5)

meanftemp <- apply(allfspec[,,targetelectrodes2,],c(1,2,4),mean,na.rm=TRUE)
meanfspec <- apply(abs(meanftemp),c(2,3),median,na.rm=TRUE)
SEfspec <- apply(abs(meanftemp),c(2,3),sd)/sqrt(46)

  h1 <- readTIFF(paste0(figdir,'/head2_',titlelist[cond],'.tiff'))
  e3 <- readTIFF('Figures/colourbar.tiff')

polygon(f[c(3:100,100:3)],c(meanfspec[cond,3:100]+SEfspec[cond,3:100],meanfspec[cond,100:3]-SEfspec[cond,100:3]),col=addalpha('cornflowerblue',0.25),border=NA)
lines(f[3:100],meanfspec[cond,3:100],lwd=3) 
# lines(c(0,50), meanfspec[1,c(53,53)],lty=2)
# lines(c(26,26),c(0,0.5),col='red')
}
dev.off()

@@

pdf('Figures/illusiondata.pdf', width = 12, height = 12)
#visualise data
  ggplot(subjecitve_data_illusion_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Illusion Score")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               c("UV.Illusion.Median","MS.Illusion.Median")),
              map_signif_level = TRUE,annotations="***",y_position = c(115,135,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#run one way RMANOVA for illusion data
#res.aov1 <- anova_test(data = subjecitve_data_illusion_1, dv = Score, wid = PID, within = Condition)
#get_anova_table(res.aov1)

#get medians and sds for illusion data
illusionmsds <- group_by(subjecitve_data_illusion_1, Condition) %>%
  summarise(
    count = n(),
    median = median(Score, na.rm = TRUE),
    sd = sd(Score, na.rm = TRUE)
  )
illusionmsds

#posthoc tests
#pwc1 <- subjecitve_data_illusion_1 %>%
  #pairwise_t_test(
    #Score ~ Condition, paired = TRUE,
    #p.adjust.method = "bonferroni"
  #)
#pwc1

@@

#posthoc tests
nppwc_illusion <- subjecitve_data_illusion_1 %>%
  wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "holm")

@@

pdf('Figures/disownershipdata.pdf', width = 12, height = 11)
#visualise data
  ggplot(subjecitve_data_disownership_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA)+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Disownerhsip Score")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  geom_signif(data=subjecitve_data_disownership_1,comparisons=list(c("NI.Disownership.Median","UV.Disownership.Median"),
                                                               c("NIT.Disownership.Median","UV.Disownership.Median"),
                                                               c("MS.Disowenership.Median","UV.Disownership.Median")),
              map_signif_level = TRUE,annotations=c("*","***","*"),y_position = c(115,130,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#posthoc tests
#pwc2 <- subjecitve_data_disownership_1 %>%
  #pairwise_t_test(
    #Score ~ Condition, paired = TRUE,
    #p.adjust.method = "bonferroni"
  #)
#pwc2

@@

#posthoc tests
nppwc_disownership <- subjecitve_data_disownership_1 %>%
  wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "holm")

@@

pdf('Figures/controldata.pdf', width = 12, height = 11)
#visualise data
  ggplot(subjecitve_data_control_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA)+
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Control Score")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#non-parametric tests
res.fried_control <- subjecitve_data_control_1 %>% friedman_test(Score ~ Condition |PID)
effectsize_control <- subjecitve_data_control_1 %>% friedman_effsize(Score ~ Condition |PID)

#posthoc tests
nppwc_control <- subjecitve_data_control_1 %>%
  wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "bonferroni")

@@

pdf('Figures/correlationdata.pdf', width = 12, height = 11)
  ggscatter(correlation_data, x = "Score", y = "Amplitude", 
          facet.by = "Condition",
          panel.labs = list( Condition=c( "NI","NIT","MS","UV") ),
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "spearman",
          cor.coef.size = 8,
          xlab = "Subjecitve Illusion Score", ylab = "SSEP Amplitude (µV)") +
    theme(text = element_text(size = 20))

@@

Resizing illusions, delivered using augmented reality, resize a body part through either stretching or shrinking manipulations. These resizing illusions have been investigated in visuotactile, visual-only, and visuo-auditory presentations. However, the neural underpinnings of these resizing illusions remain undefined. This study sought to understand the neural mechanisms behind these illusions (n = 46), by using somatosensory steady state evoked potentials in addition to subjective self-report questionnaires, to enhance knowledge of what drives the subjective embodiment during resizing illusions. Since these Illusions have been shown to provide analgesic effects for individuals with chronic pain conditions, this study also aimed to provide an empirical basis for future investigations in chronic pain samples undergoing resizing illusions. Confirmatory results demonstrated significant differences in subjective experience between non-illusion and multisensory illusion conditions, and EEG data measuring somatosensory response across electrodes of interest (F1 & FC1) to 26Hz stimulation to the resized digit showed a significant effect of condition with reduced amplitudes in illusory conditions compared to non-illusion conditions. Exploratory analysis of electrodes with the greatest SSEP response across conditions (FPZ & FCZ) found no effect of condition but did show reduced amplitudes in illusory and non-illusion tactile conditions compared to the non-illusion condition. These findings demonstrate a potential sharpening of neural representations as a result of both illusory stretching and tactile input in participants without chronic pain, thereby providing a basis for investigations of comparable subjective and steady state illusion responses in a chronic pain population, who are thought to have more diffuse neural representations of their affected body parts.

@@

Using different sensory manipulations of finger resizing illusions, in addition to using an electromagnetic solenoid stimulator, this study aimed to investigate subjective illusory experience and SSEP responses in participants without chronic pain, to better understand the experience of body ownership illusions from subjective experience and cortical representation perspectives. To test this, different finger resizing illusions consisting of multisensory (visuotactile) stretching (MS), unimodal-visual stretching (UV), a non-illusion control condition without tactile input (NI), and a non-illusion control condition with tactile input (NIT) were used to assess alternate aspects of illusory resizing manipulations and their related effects on SSEP response. The inclusion of two control conditions (NI, NIT) was to assess whether localisation of cortical representations arise from resizing manipulations to the finger, or from tactile input given to the finger. The first hypothesis, acting as a positive control (1), was that there will be a greater illusory experience, measured via a subjective illusory experience questionnaire, in the (1a) MS condition compared to the NI condition and in the (1b) MS condition compared to the NIT condition. The main experimental hypothesis for this study was that (2) there will be a significant difference in SSEP response across the electrodes of interest (F1 & FC1, see section 3. Pilot Data) when comparing across all conditions.  Subsequent hypotheses were that there would be significant differences in SSEP response when comparing (2a) MS visuotactile illusory resizing to the NI condition, when comparing (2b) UV illusory resizing to the NI condition, but (2c) that there would be no significant difference when comparing the NIT condition to the NI condition.

@@

The subjective illusory experience questionnaire was  used as a positive control for the current study. Previous research has shown significantly greater illusion strength for MS conditions compared to non-illusion conditions (Carey et al., 2019; Hansford et al., 2023a), which we attempted to replicate. Questionnaire data were be analysed using R (R Core Team, 2021). A one-way ANOVA (or non-parametric equivalent)would be run to compare the dependent variable of normalised (baseline corrected) illusion score from each independent condition. Given significant findings, post-hoc tests would be run, with Bonferroni correction for 2 comparisons (MS Vs NI, MS Vs NIT) at an initial alpha of 0.05.

@@

*There will be a significant difference in SSEP response across the electrodes of interest (F1 & FC1) when comparing (2a) the MS condition to the NI condition, when comparing (2b) the UV condition to the NI condition, but (2c) that there will be no significant difference when comparing the NIT condition to the NI condition.*

>>>>

---
title: "Illusory Finger Stretching and Somatosensory Responses"
author: ""
bibliography: references-rr.bib
#csl: elife.csl
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: true
execute: 
  echo: false
  include: false
  output: false
format: 
  pdf:
    fig-pos: "H"
---

@@

# install R packages
packagelist <- c('tidyverse','ggpubr','ez','ggplot2','jpeg','grid','plotrix','lavaan', 'cowplot', 'ggpubr','rstatix','reticulate','moments','tiff','rmarkdown','tinytex','osfr','FSA')  
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

pdf('Figures/eegdata1.pdf', width = 12, height = 11)
#visualise data
  ggplot(eegdata_2, aes(x = Condition, y = Amplitude)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="black", fill="black") +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("SSEP Amplitude (µV)")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_y_log10() +
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  #geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               #c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               #c("UV.Illusion.Median","MS.Illusion.Median")),
              #map_signif_level = TRUE,annotations="***",y_position = c(115,135,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

pdf('Figures/eegdata2.pdf', width = 12, height = 11)
#visualise data
  ggplot(eegdata_exp2, aes(x = Condition, y = Amplitude)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="black", fill="black") +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("SSEP Amplitude (µV)")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_y_log10() +
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  #geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               #c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               #c("UV.Illusion.Median","MS.Illusion.Median")),
              #map_signif_level = TRUE,annotations="***",y_position = c(115,135,145))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#get medians and sds for eeg data
eegdatamsds <- group_by(eegdata_2, Condition) %>%
  summarise(
    count = n(),
    median = median(Amplitude, na.rm = TRUE),
    mean = mean(Amplitude, na.rm = TRUE),
    sd = sd(Amplitude, na.rm = TRUE)
  )

@@

#run one way RMANOVA for illusion data
res.aov4 <- anova_test(data = eegdata_2, dv = Amplitude, wid = PID, within = Condition, effect.size = "pes")
get_anova_table(res.aov4)

#posthoc tests
pwc4 <- eegdata_2 %>%
pairwise_t_test(
  Amplitude ~ Condition, paired = TRUE,
  p.adjust.method = "bonferroni",
  detailed = TRUE
)
pwc4

@@

#posthoc tests
nppwc_eeg1 <- eegdata_2 %>%
  wilcox_test(Amplitude ~ Condition, paired = TRUE, comparisons = list(c("V1", "V3"), c("V1", "V4"), c("V1", "V2")), p.adjust.method = "holm",
    detailed = TRUE)

nppwc2_eeg2 <- eegdata_exp2 %>%
  wilcox_test(Amplitude ~ Condition, paired = TRUE, comparisons = list(c("V1", "V3"), c("V1", "V4"), c("V1", "V2")), p.adjust.method = "holm",
    detailed = TRUE)

@@

#for individual plots uncomment next line
#for (subj in 1:46){

#create spectra plots
#for individual plots comment next line
pdf(paste0(figdir,"/spectrum1.pdf"), width=12, height=12)
#for individual plots uncomment next line
#pdf(paste0(figdir,"/spectrum",subj,".pdf"), width=12, height=12)
par(mfrow=c(2,2))
# plot the Fourier spectra across all electrodes
f <- seq(0,499.5,0.5)

@@

  #for individual plots comment next line
polygon(f[c(3:100,100:3)],c(meanfspec[cond,3:100]+SEfspec[cond,3:100],meanfspec[cond,100:3]-SEfspec[cond,100:3]),col=addalpha('cornflowerblue',0.25),border=NA)
  #for individual plots comment next line
lines(f[3:100],meanfspec[cond,3:100],lwd=3) 
  #for individual plots uncomment next line
#lines(f[3:100],abs(meanftemp[subj,cond,3:100]),lwd=1)
# lines(c(0,50), meanfspec[1,c(53,53)],lty=2)
# lines(c(26,26),c(0,0.5),col='red')
}
dev.off()
#for individual plots uncomment next line
#}

#create exploratory spectra plots
# pdf(paste0(figdir,"/spectrum_other.pdf"), width=12, height=12)
# par(mfrow=c(2,2))
# # plot the Fourier spectra across all electrodes
# f <- seq(0,499.5,0.5)
# 
#   ticklocsx <- seq(0,50,10)
#   ticklocsy <- seq(0,0.5,0.1)
# 
# meanftemp <- apply(allfspec[,,targetelectrodes2,],c(1,2,4),mean,na.rm=TRUE)
# meanfspec <- apply(abs(meanftemp),c(2,3),median,na.rm=TRUE)
# SEfspec <- apply(abs(meanftemp),c(2,3),sd)/sqrt(46)
# 
# for (cond in 1:4){
#   
#   h1 <- readTIFF(paste0(figdir,'/head2_',titlelist[cond],'.tiff'))
#   e3 <- readTIFF('Figures/colourbar.tiff')
# 
#   plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(min(ticklocsx),max(ticklocsx)), ylim=c(min(ticklocsy),max(ticklocsy)))
#   axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)     # plot tick marks (no labels)
#   axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
#   mtext(text = ticklocsx, side = 1, at=ticklocsx, line=0.5, cex=1.3)
#   mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1, cex=1.3)
#   title(xlab='Frequency (Hz)', col.lab=rgb(0,0,0), line=2, cex.lab=2.5)
#   title(ylab='Amplitude (µV)', col.lab=rgb(0,0,0), line=2.4, cex.lab=2.5)
#   title(main=titlelist[cond],cex.main=2)  
# 
#   aspratio <- 90  # this is the aspect ratio of the output pdf
#   imwidth <- 0.4
#   xstart <- 18
#   ystart <- 0.15
#   rasterImage(h1,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth) 
#   
#   imwidth <- 0.1
#   xstart <- 40
#   ystart <- 0.12
#   rasterImage(e3,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)
#   
#   text(38,0.12,'0',pos=4,cex=1.8)
#   text(34,0.2,'1.15',pos=4,cex=1.8)
#   text(49,0.24,'Amplitude',pos=4,cex=2,srt=270)
#   
# # -- polygon(f[c(3:100,100:3)],c(meanfspec[cond,3:100]+SEfspec[cond,3:100],meanfspec[cond,100:3]-SEfspec[cond,100:3]),col=addalpha('cornflowerblue',0.25),border=NA)
# 
# # - -lines(f[3:100],meanfspec[cond,3:100],lwd=3) 
# # lines(c(0,50), meanfspec[1,c(53,53)],lty=2)
# # lines(c(26,26),c(0,0.5),col='red')
# }
# dev.off()

@@

pdf('Figures/illusiondata.pdf', width = 12, height = 12)
#visualise data
  ggplot(subjecitve_data_illusion_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA, color = "black")+
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="black", fill="black") +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Illusion Score")+
  ylim(-100,138) +
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  geom_signif(data=subjecitve_data_illusion_1,comparisons=list(c("NI.Illusion.Median","MS.Illusion.Median"),
                                                               c("NIT.Illusion.Median","MS.Illusion.Median"),
                                                               c("UV.Illusion.Median","MS.Illusion.Median")),
              map_signif_level = TRUE,annotations="***",y_position = c(105,120,129))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#run one way RMANOVA for illusion data
res.aov1 <- anova_test(data = subjecitve_data_illusion_1, dv = Score, wid = PID, within = Condition,effect.size = "pes")
get_anova_table(res.aov1)

#get medians and sds for illusion data
illusionmsds <- group_by(subjecitve_data_illusion_1, Condition) %>%
  summarise(
    count = n(),
    median = median(Score, na.rm = TRUE),
    mean = mean(Score, na.rm = TRUE),
    sd = sd(Score, na.rm = TRUE)
  )
illusionmsds

#posthoc tests
pwc1 <- subjecitve_data_illusion_1 %>%
  pairwise_t_test(
    Score ~ Condition, paired = TRUE,
    p.adjust.method = "bonferroni",
    detailed = TRUE
  )
pwc1

@@

#posthoc tests
nppwc_illusion <- subjecitve_data_illusion_1 %>%
  wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "holm",
    detailed = TRUE)

@@

pdf('Figures/disownershipdata.pdf', width = 12, height = 11)
#visualise data
  ggplot(subjecitve_data_disownership_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA)+
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="black", fill="black") +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Disownerhsip Score")+
  ylim(-100,138) +
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  geom_signif(data=subjecitve_data_disownership_1,comparisons=list(c("NI.Disownership.Median","UV.Disownership.Median"),
                                                               c("NIT.Disownership.Median","UV.Disownership.Median"),
                                                               c("MS.Disowenership.Median","UV.Disownership.Median")),
              map_signif_level = TRUE,annotations=c("*","***","*"),y_position = c(105,120,129))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#posthoc tests
# pwc2 <- subjecitve_data_disownership_1 %>%
#   pairwise_t_test(
#     Score ~ Condition, paired = TRUE,
#     p.adjust.method = "bonferroni",
#     detailed = TRUE
#   )

@@

#posthoc tests
nppwc_disownership <- subjecitve_data_disownership_1 %>%
  wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "holm",
    detailed = TRUE)

@@

pdf('Figures/controldata.pdf', width = 12, height = 11)
#visualise data
  ggplot(subjecitve_data_control_1, aes(x = Condition, y = Score)) + 
  geom_violin(trim = FALSE) + 
  geom_violin(aes(color = Condition), trim = FALSE) + 
  geom_boxplot(width = 0.2, outlier.shape = NA)+
  stat_summary(fun.y=mean, geom="point", shape=20, size=5, color="black", fill="black") +
  geom_dotplot(binaxis = "y",
               stackdir = "center",
               dotsize = 0.5,
               alpha = 0.2) +
  ylab("Control Score")+
  font("ylab", size = 20)+
  font("xlab", size = 20)+
  font("xy.text", size = 15)+
  scale_x_discrete(labels = c('NI','NIT','MS','UV'))+
  scale_color_manual(values = c("cornflowerblue", "cornflowerblue","cornflowerblue", "cornflowerblue"))+
  theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),legend.position = "none", panel.background = element_blank(),axis.line = element_line(colour = "black"))
dev.off()

@@

#non-parametric tests
kruskall.wallis <- kruskal.test(Score ~ Condition, data = subjecitve_data_control_1)
#res.fried_control <- subjecitve_data_control_1 %>% friedman_test(Score ~ Condition |PID)
#effectsize_control <- subjecitve_data_control_1 %>% friedman_effsize(Score ~ Condition |PID)

#posthoc tests
# nppwc_control <- subjecitve_data_control_1 %>%
#   wilcox_test(Score ~ Condition, paired = TRUE, p.adjust.method = "bonferroni",
#     detailed = TRUE)

nppwc_controlD <- dunnTest(Score ~ Condition,
         data=subjecitve_data_control_1,
         method="bonferroni")

@@

pdf('Figures/correlationdata.pdf', width = 12, height = 11)
  ggscatter(correlation_data, x = "Score", y = "Amplitude", 
          facet.by = "Condition",
          panel.labs = list( Condition=c( "NI","NIT","MS","UV") ),
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "spearman",
          cor.coef.size = 8,
          xlab = "Subjecitve Illusion Score", ylab = "SSEP Amplitude (µV)") +
    theme(axis.text.x = element_text(size = 20), axis.text.y = element_text(size = 20),axis.title.x = element_text(size = 20), axis.title.y = element_text(size = 20),text = element_text(size = 20))

@@

Resizing illusions, delivered using augmented reality, resize a body part through either stretching or shrinking manipulations. These resizing illusions have been investigated in visuotactile, visual-only, and visuo-auditory presentations. However, the neural underpinnings of these resizing illusions remain undefined. This study sought to understand the neural mechanisms behind these illusions (n = 46), by using somatosensory steady state evoked potentials in addition to subjective self-report questionnaires, to enhance knowledge of what drives the subjective embodiment during resizing illusions. Since these Illusions have been shown to provide analgesic effects for individuals with chronic pain conditions, this study also aimed to provide an empirical basis for future investigations in chronic pain samples undergoing resizing illusions. Confirmatory results demonstrated significant differences in subjective experience between non-illusion and multisensory illusion conditions, whilst EEG data measuring somatosensory response across electrodes of interest (F1 & FC1) to 26Hz stimulation to the resized digit showed no significant effects of condition. Exploratory non-parametric SSEP analyses showed a significant effect of condition, with reduced amplitudes in illusory conditions compared to non-illusion conditions, however no significant differences arose in exploratory post hoc tests. These findings demonstrate no clear effect of resizing illusions on somatosensory SSEP amplitudes, but could hint to a potential sharpening of neural representations as a result of illusory stretching in participants without chronic pain, thereby providing a basis for investigations of comparable subjective and steady state illusion responses in a chronic pain population, who are thought to have more diffuse neural representations of their affected body parts.

@@

Using different sensory manipulations of finger resizing illusions, in addition to using an electromagnetic solenoid stimulator, this study aimed to investigate subjective illusory experience and somatosensory SSEP responses in participants without chronic pain, to better understand the experience of body ownership illusions from subjective experience and cortical representation perspectives. To test this, different finger resizing illusions consisting of multisensory (visuotactile) stretching (MS), unimodal-visual stretching (UV), a non-illusion control condition without tactile input (NI), and a non-illusion control condition with tactile input (NIT) were used to assess alternate aspects of illusory resizing manipulations and their related effects on somatosensory SSEP response. The inclusion of two control conditions (NI, NIT) was to assess whether localisation of cortical representations arise from resizing manipulations to the finger, or from tactile input given to the finger. The first hypothesis, acting as a positive control (1), was that there will be a greater illusory experience, measured via a subjective illusory experience questionnaire, in the (1a) MS condition compared to the NI condition and in the (1b) MS condition compared to the NIT condition. The main experimental hypothesis for this study was that (2) there will be a significant difference in somatosensory SSEP response across the electrodes of interest (F1 & FC1, see Appendix A Pilot Data) when comparing across all conditions.  Subsequent hypotheses were that there would be significant differences in SSEP response when comparing (2a) the MS condition to the NI condition, when comparing (2b) the UV condition to the NI condition, but (2c) that there would be no significant difference when comparing the NIT condition to the NI condition.

@@

The subjective illusory experience questionnaire was  used as a positive control for the current study. Previous research has shown significantly greater illusion strength for MS conditions compared to non-illusion conditions (Carey et al., 2019; Hansford et al., 2023a), which we attempted to replicate. Questionnaire data were be analysed using R (R Core Team, 2021). A one-way ANOVA would be run to compare the dependent variable of normalised (control indexed) illusion score from each independent condition. Given significant findings, post-hoc tests would be run, with Bonferroni correction for 2 comparisons (MS Vs NI, MS Vs NIT) at an initial alpha of 0.05.

@@

*(2) There will be a significant difference in somatosensory SSEP response across the electrodes of interest (F1 & FC1). Subsequent hypotheses are that there will be a significant difference in SSEP response in (2a) the MS condition compared to the NI condition, and (2b) the UV condition compared to the NI condition, but (2c) that there will be no significant difference when comparing the NIT condition to the NI condition. *
